---
title: "API work"
author: "Haoheng Tang"
date: "3/2/2021"
output: rmarkdown::github_document
---
## 1. Setup

### 1.1 Load packages

```{r load packages, include=TRUE, echo=FALSE, message=FALSE}
library(tidyverse)
library(sf)
library(geojsonsf)
library(QuantPsyc)
library(RSocrata)
library(viridis)
library(caret)
library(spatstat)
library(spdep)
library(FNN)
library(grid)
library(gridExtra)
library(knitr)
library(kableExtra)
library(tidycensus)

library(mapview)
library(httr)
library(dplyr)
library(readxl)
library(stringr)
```

### 1.2 Load fire data

```{r load fire data, include=TRUE, echo=FALSE, message=FALSE}
library(readxl)
fire <- read_excel("./data/fire.xlsx")
library(stringr)

fire1 = fire %>%
  filter(addr_type ==1)
fire1$address <- paste(ifelse(is.na(fire1$number)==FALSE,fire1$number,''),
                       "%20",
                       ifelse(is.na(fire1$st_prefix)==FALSE,fire1$st_prefix,''),
                       ifelse(is.na(fire1$st_prefix)==FALSE,"%20",''),
                       ifelse(is.na(fire1$street)==FALSE,fire1$street,''),
                       "%20",
                       ifelse(is.na(fire1$st_type)==FALSE,fire1$st_type,''), sep = "")
fire2 = fire %>%
  filter(addr_type ==2)
fire2$address <- paste(ifelse(is.na(fire2$xst_prefix)==FALSE,fire2$xst_prefix,''),
                       ifelse(fire2$xst_prefix!='',"%20",''),
                       ifelse(is.na(fire2$xstreet)==FALSE,fire2$xstreet,''),
                       "%20",
                       ifelse(is.na(fire2$xst_type)==FALSE,fire2$xst_type,''),
                       "%20",
                       "&",
                       "%20",
                       ifelse(is.na(fire2$st_prefix)==FALSE,fire2$st_prefix,''), 
                       ifelse(fire2$st_prefix!='',"%20",''),
                       ifelse(is.na(fire2$street)==FALSE,fire2$street,''), 
                       "%20",
                       ifelse(is.na(fire2$st_type)==FALSE,fire2$st_type,''),
                       sep = "")  
```


## 2. Request opa_account_num & parcel_id & census_tract&block

```{r opa_account_num & parcel_id, include=TRUE, echo=TRUE, message=FALSE}
fireData <- rbind(fire1, fire2) %>%
  dplyr::select(address)

fireData$MUSA_ID <- paste0("MUSA_",1:nrow(fireData))

for (i in 1:50) {
  address  <- fireData$address[[i]]
  base_url <- "http://api.phila.gov/ais/v1/"
  endpoint <- "search/"
  key      <- "?gatekeeperKey=dc953bbc4ade9d00eabf5409f6d73d3e"
  url <- paste(base_url, endpoint, address, key, sep="")
  response <- httr::GET(url)
  tidy_res <- httr::content(response, simplifyVector=TRUE)
  if (length(tidy_res) != 4){
    if(length(tidy_res$features$properties$opa_account_num)==2)
      opa_num <-  tidy_res$features$properties$opa_account_num[2]
    else
      opa_num <-  tidy_res$features$properties$opa_account_num[1]
    if(is.null(opa_num)){
      opa_num <- "OPA IS NULL"
    } else if(nchar(opa_num)==0) {
      opa_num <- "OPA IS ZERO LENGTH"
    }
    cat("Address",i,"OPA#:",opa_num,"\n")
    fireData[i,"opa_account_num"] <- opa_num
  }else{
    cat("Address",i,"NO ADDRESS FOUND!","\n")
    fireData[i,"opa_account_num"] <- "NONE FOUND"
  }
  
  if (length(tidy_res) != 4){
    if(length(tidy_res$features$properties$dor_parcel_id)==2)
      parcel_id <-  tidy_res$features$properties$dor_parcel_id[2]
    else
      parcel_id <-  tidy_res$features$properties$dor_parcel_id[1]
    if(is.null(parcel_id)){
      parcel_id <- "PARCEL_ID IS NULL"
    } else if(nchar(parcel_id)==0) {
      parcel_id <- "0LENGTH"
    }
    cat("Address",i,"parcel_id#:",parcel_id,"\n")
    fireData[i,"parcel_id"] <- parcel_id
  }else{
    cat("Address",i,"NO ADDRESS FOUND!","\n")
    fireData[i,"parcel_id"] <- "NONE FOUND"
  }
  
  if(length(tidy_res) != 4){
    if(length(tidy_res$features$properties$census_tract_2010)==2){
      census_tract <-  tidy_res$features$properties$census_tract_2010[2]}
    else{
      census_tract <- tidy_res$features$properties$census_tract_2010
    }
    if(length(tidy_res$features$properties$census_block_2010)==2){
      census_block <-  tidy_res$features$properties$census_block_2010[2]}
    else{
      census_block <- tidy_res$features$properties$census_block_2010
    }
    fireData[i,"census_tract"] <- as.character(census_tract)
    cat("Address",i,"census_tract#:",census_tract,"\n")
    fireData[i,"census_block"] <- as.character(census_block)
    cat("Address",i,"census_block#:",census_block,"\n")
  }else{
    cat("Address",i,"NO ADDRESS FOUND!","\n")
    fireData[i,"census_tract"] <- "NONE FOUND"
    fireData[i,"census_block"] <- "NONE FOUND"
  }
}

fireData <- fireData %>%
  mutate(parcel_id = ifelse(str_count(parcel_id)==10,parcel_id,substr(parcel_id,1,10)))
```

```{r 2020 fire data, include=TRUE, echo=TRUE, message=FALSE}
fire2020 <- read_excel("./data/2020 Fires for U of P.xlsx")

fire2020$address <- paste(ifelse(is.na(fire2020$Number)==FALSE,fire2020$Number,''),
                       "%20",
                       ifelse(is.na(fire2020$St_prefix)==FALSE,fire2020$St_prefix,''),
                       ifelse(is.na(fire2020$St_prefix)==FALSE,"%20",''),
                       ifelse(is.na(fire2020$Street)==FALSE,fire2020$Street,''),
                       "%20",
                       ifelse(is.na(fire2020$St_type)==FALSE,fire2020$St_type,''), 
                       sep = "")
```


```{r request 2020 opa&parcel_id,include=TRUE, echo=TRUE, message=FALSE}
for (i in 3301:nrow(fire2020)) {
  address  <- fire2020$address[[i]]
  base_url <- "http://api.phila.gov/ais/v1/"
  endpoint <- "search/"
  key      <- "?gatekeeperKey=dc953bbc4ade9d00eabf5409f6d73d3e"
  url <- paste(base_url, endpoint, address, key, sep="")
  response <- httr::GET(url)
  tidy_res <- httr::content(response, simplifyVector=TRUE)
  if (length(tidy_res) != 4){
    if(length(tidy_res$features$properties$opa_account_num)==2)
      opa_num <-  tidy_res$features$properties$opa_account_num[2]
    else
      opa_num <-  tidy_res$features$properties$opa_account_num[1]
    if(is.null(opa_num)){
      opa_num <- "OPA IS NULL"
    } else if(nchar(opa_num)==0) {
      opa_num <- "OPA IS ZERO LENGTH"
    }
    cat("Address",i,"OPA#:",opa_num,"\n")
    fire2020[i,"opa_account_num"] <- opa_num
  }else{
    cat("Address",i,"NO ADDRESS FOUND!","\n")
    fire2020[i,"opa_account_num"] <- "NONE FOUND"
  }
  
  if (length(tidy_res) != 4){
    if(length(tidy_res$features$properties$dor_parcel_id)==2)
      parcel_id <-  tidy_res$features$properties$dor_parcel_id[2]
    else
      parcel_id <-  tidy_res$features$properties$dor_parcel_id[1]
    if(is.null(parcel_id)){
      parcel_id <- "PARCEL_ID IS NULL"
    } else if(nchar(parcel_id)==0) {
      parcel_id <- "0LENGTH"
    }
    cat("Address",i,"parcel_id#:",parcel_id,"\n")
    fire2020[i,"parcel_id"] <- parcel_id
  }else{
    cat("Address",i,"NO ADDRESS FOUND!","\n")
    fire2020[i,"parcel_id"] <- "NONE FOUND"
  }
}
```

```{r}
allfire <- read.csv("all_opa_par_addr.csv")

allfire$MUSA_ID <- paste0("MUSA_",1:nrow(allfire))

write.csv(allfire, "new_opa_par_addr.csv")
```



```{r combine 2020 with previous fireData,include=TRUE, echo=TRUE, message=FALSE}
fire2020$MUSA_ID <- paste0("MUSA_",1:nrow(fire2020),"_2020")

opa_par_addr <- read_csv("opa_par_addr.csv") %>%
  dplyr::select(-X1)

fire2020 <- fire2020 %>%
  rename(Parcel_Id = parcel_id)

opa_par_addr <- opa_par_addr %>%
  rbind(fire2020 %>%
          dplyr::select(opa_account_num, Parcel_Id, address, MUSA_ID))

write.csv(fire2020, "fire2020_opa.csv")
write.csv(opa_par_addr, "all_opa_par_addr.csv")
```


```{r request 2020 census_tract&block numbers,include=TRUE, echo=TRUE, message=FALSE}
for (i in 1:300) {
  address  <- fire2020$address[[i]]
  base_url <- "http://api.phila.gov/ais/v1/"
  endpoint <- "search/"
  key      <- "?gatekeeperKey=dc953bbc4ade9d00eabf5409f6d73d3e"
  url <- paste(base_url, endpoint, address, key, sep="")
  response <- httr::GET(url)
  tidy_res <- httr::content(response, simplifyVector=TRUE)  
  if(length(tidy_res) != 4){
    if(length(tidy_res$features$properties$census_tract_2010)==2){
      census_tract <-  tidy_res$features$properties$census_tract_2010[2]}
    else{
      census_tract <- tidy_res$features$properties$census_tract_2010
    }
    if(length(tidy_res$features$properties$census_block_2010)==2){
      census_block <-  tidy_res$features$properties$census_block_2010[2]}
    else{
      census_block <- tidy_res$features$properties$census_block_2010
    }
    fire2020[i,"census_tract"] <- as.character(census_tract)
    cat("Address",i,"census_tract#:",census_tract,"\n")
    fire2020[i,"census_block"] <- as.character(census_block)
    cat("Address",i,"census_block#:",census_block,"\n")
  }else{
    cat("Address",i,"NO ADDRESS FOUND!","\n")
    fire2020[i,"census_tract"] <- "NONE FOUND"
    fire2020[i,"census_block"] <- "NONE FOUND"
  }
}

fire2020 <- fire2020 %>%
  mutate(parcel_id = ifelse(str_count(parcel_id)==10,parcel_id,substr(parcel_id,1,10)))
```


## 3. Request potential feature data

### 3.1 Parcel geometry

```{r parcel, include=TRUE, echo=TRUE, message=FALSE}
DOR_Parcel <- fireData[1:50,]

for (i in 1:50) {
  if(DOR_Parcel$parcel_id[[i]] != "0LENGTH" && DOR_Parcel$parcel_id[[i]] != "NONE FOUND" &&DOR_Parcel$parcel_id[[i]] != "PARCEL_ID IS NULL"){
      base <- "https://services.arcgis.com/fLeGjb7u4uXqeF9q/arcgis/rest/services/DOR_Parcel/FeatureServer/0/query?outFields=*&where=BASEREG%3D%27"
      BASEREG <- DOR_Parcel$parcel_id[[i]]
      end <- "%27&returnCentroid=true&f=pjson"
      dor_url <- paste(base, BASEREG, end, sep="")
      get_dor <- httr::GET(dor_url)
      dor_data <- httr::content(get_dor, simplifyVector=TRUE, as = "parsed", type = "application/json")
      DOR_Parcel[i,"Parcel_OBJECTID"] <- dor_data$features$attributes$OBJECTID
      DOR_Parcel[i,"PARCEL"] <- dor_data$features$attributes$PARCEL
      DOR_Parcel[i,"Shape__Area"] <- dor_data$features$attributes$Shape__Area
      DOR_Parcel[i,"Shape__Length"] <- dor_data$features$attributes$Shape__Length
      DOR_Parcel[i,"ADDR_SOURCE"] <- dor_data$features$attributes$ADDR_SOURCE
      DOR_Parcel[i,"x"] <- dor_data$features$centroid$x
      DOR_Parcel[i,"y"] <- dor_data$features$centroid$y
      #DOR_Parcel[i,"geometry"] <- dor_data$features$geometry
  }

}

DOR_Parcel.sf <- DOR_Parcel %>%
  drop_na(x)%>%
  st_as_sf(coords = c("x","y"), crs = 3857)

```

### 3.2 L&I violation

```{r L&I violation, include=TRUE, echo=TRUE, message=FALSE}
violation <- fireData[1:50,]

for (i in 1:50) {
  base_url <- "https://phl.carto.com/api/v2/"
  endpoint <- "sql"
  query    <- c("?q=SELECT%20*%20FROM%20violations%20WHERE%20opa_account_num%20=%20")
  opa_num  <- paste0("%27",fireData$opa_account_num[[i]],"%27")
  url <- paste(base_url, endpoint, query, opa_num, sep="")
  response <- httr::GET(url)
  tidy_res <- httr::content(response, simplifyVector=TRUE)
  
  if (response$status_code != 400){
    if(length(tidy_res$rows$violationcode)==1){
    vio_code <-  tidy_res$rows$violationcode
    vio_title <- tidy_res$rows$violationcodetitle
  
    violation$vio_code[[i]] <- vio_code
    violation$vio_title[[i]] <- vio_title
    
    cat("Address",i,vio_code, vio_title, "\n")}
    else{
    violation$vio_code[[i]] <- "NO CODE VIOLATION"
    violation$vio_title[[i]] <- "NO CODE VIOLATION"
    cat("Address",i,"NO CODE VIOLATION\n")
    }
  }
  else{
    violation$vio_code[[i]] <- "NO RESPONSE"
    violation$vio_title[[i]] <- "NO RESPONSE"
    cat("Address",i,"NO RESPONSE\n")
  }
}
```

### 3.3 Properties

```{r Properties, include=TRUE, echo=TRUE, message=FALSE}
#"https://phl.carto.com/api/v2/sql?q=SELECT * FROM opa_properties_public"

properties <- fireData[1:50,]

for (i in 1:50) {
  base_url <- "https://phl.carto.com/api/v2/"
  endpoint <- "sql"
  query    <- c("?q=SELECT%20*%20FROM%20opa_properties_public%20WHERE%20parcel_number%20=%20")
  opa_num  <- paste0("%27",properties$opa_account_num[[i]],"%27")
  prop_url <- paste(base_url, endpoint, query, opa_num, sep="")
  response_prop <- httr::GET(prop_url)
  tidy_res_prop <- httr::content(response_prop, simplifyVector=TRUE)
  
  if (response_prop$status_code != 400){
    total_area <-  tidy_res_prop$rows$total_area
    total_livable_area <- tidy_res_prop$rows$total_livable_area
    zoning <- tidy_res_prop$rows$zoning
    category_code <- tidy_res_prop$rows$category_code
    category <- case_when(category_code == 1 ~ "Residential",
                          category_code == 2 ~ "Hotels and Apartments",
                          category_code == 3 ~ "Store with Dwelling",
                          category_code == 4 ~ "Commercial",
                          category_code == 5 ~ "Industrial",
                          category_code == 6 ~ "Vacant Land")
    interior_condition <- tidy_res_prop$rows$interior_condition
    interior <- case_when(interior_condition == 0 ~ "Not Applicable",
                          interior_condition == 2 ~ "New/Rehabbed",
                          interior_condition == 3 ~ "Above Average",
                          interior_condition == 4 ~ "Average",
                          interior_condition == 5 ~ "Below Average",
                          interior_condition == 6 ~ "Vacant",
                          interior_condition == 7 ~ "Sealed/Structurally Compromised")
  
    properties$total_area[[i]] <- total_area
    properties$total_livable_area[[i]] <- total_livable_area
    properties$zoning[[i]] <- zoning
    properties$category[[i]] <- category
    properties$interior[[i]] <- interior
    
    cat("\nAddress",i,"total_area", total_area)
    cat("   total_livable_area", total_livable_area)
    cat("   zoning", zoning)
    cat("   category", category)
    cat("   interior", interior)
    }
  else{
    properties$total_area[[i]] <- "NO RESPONSE"
    properties$total_livable_area[[i]] <- "NO RESPONSE"
    properties$zoning[[i]] <- "NO RESPONSE"
    properties$category[[i]] <- "NO RESPONSE"
    properties$interior[[i]] <- "NO RESPONSE"
    cat("\nAddress",i,"NO RESPONSE")
  }
}
```

### 3.4 311 request

#### draw boundaries

```{r calculate distance, include=TRUE, echo=TRUE, message=FALSE}
DOR_4326 <- DOR_Parcel.sf %>% 
  st_transform(crs = 4326)

distance <- 100
DOR_meters <- DOR_4326 %>%  
  st_transform(32618) %>% 
  cbind(st_coordinates(.)) %>% 
  mutate(Xmin = X - distance,
         Xmax = X + distance,
         Ymin = Y - distance,
         Ymax = Y + distance) 

DOR_latlng <- DOR_meters %>% 
  st_drop_geometry() %>% 
  dplyr::select(X, Y, parcel_id) %>% 
  st_as_sf(coords=c("X","Y"),
           remove = FALSE,
           crs = 32618) %>% 
  st_transform(crs = 4326) %>%
  cbind(st_coordinates(.)) %>%
  rename(LNG = X.1, LAT = Y.1)

LL <- DOR_meters %>% 
  st_drop_geometry() %>% 
  dplyr::select(Xmin, Ymin, parcel_id) %>% 
  st_as_sf(coords=c("Xmin","Ymin"),
           remove = FALSE,
           crs = 32618) %>% 
  st_transform(crs = 4326) %>%
  cbind(st_coordinates(.))

UR <- DOR_meters %>% 
  st_drop_geometry() %>% 
  dplyr::select(Xmax, Ymax, parcel_id) %>% 
  st_as_sf(coords=c("Xmax","Ymax"),
           remove = FALSE,
           crs = 32618)%>% 
  st_transform(crs = 4326) %>%
  cbind(st_coordinates(.))

# mapview(DOR_4326) + mapview(LL, col.regions = "red") + mapview(UR, col.regions = "green")
```


#### Request nearby 311
```{r nearby 311 request, include=TRUE, echo=TRUE, message=FALSE}
# ref311 = "https://phl.carto.com/api/v2/sql?q=SELECT * FROM public_cases_fc WHERE requested_datetime >= '2021-03-01' AND requested_datetime < '2021-03-15' AND lat < "
#"AND lat > "
#"AND lon < "
#"AND lon > "

base311 = ("https://phl.carto.com/api/v2/sql?q=SELECT%20*%20FROM%20public_cases_fc%20WHERE%20")
where1 = paste("requested_datetime%20%3e%3d%20%27",Sys.Date()-30,
               "%27%20AND%20requested_datetime%20%3c%20%27", Sys.Date(),
               "%27%20AND%20lat%20%3C%20",sep="")
where2 = "AND%20lat%20%3E%20"
where3 = "AND%20lon%20%3C%20"
where4 = "AND%20lon%20%3E%20"

LATmax = UR$Y[[1]]
LATmin = LL$Y[[1]]
LNGmax = UR$X[[1]]
LNGmin = LL$X[[1]]

url311 <- paste(base311, where1, LATmax, where2, LATmin, where3, LNGmax, where4, LNGmin, sep="")


response311 <- httr::GET(url311)
tidy_res311 <- httr::content(response311, simplifyVector=TRUE)


if(length(tidy_res311$rows) != 0){
  request311 <- tidy_res311$rows %>%
    data.frame() %>%
    dplyr::select(service_request_id, status, service_name, service_code, requested_datetime, updated_datetime, address, lat, lon)
}else{
  request311 <- data.frame(Response=c("No 311 request within 100 meters in the last 15 days"))
}
```


#### Request 311 in specific category
```{r specific 311 request, include=TRUE, echo=TRUE, message=FALSE}
# ref311 = "https://phl.carto.com/api/v2/sql?q=SELECT * FROM public_cases_fc WHERE requested_datetime >= '2021-03-01' AND requested_datetime < '2021-03-15' AND service_name IN ('Alley Light Outage', 'No Heat (Residential)' , 'Fire Residential or Commercial' , 'Infestation Residential' ,  'Smoke Detector', 'Building Dangerous')"


"https://phl.carto.com/api/v2/sql?q=SELECT%20*%20FROM%20public_cases_fc%20WHERE%20requested_datetime%20%3E=%20%272021-03-01%27%20AND%20requested_datetime%20%3C%20%272021-03-15

%27%20AND%20service_name%20IN%20(%27Alley%20Light%20Outage%27,%20%27No%20Heat%20(Residential)%27%20,%20%27Fire%20Residential%20or%20Commercial%27%20,%20%27Infestation%20Residential%27%20,%20%20%27Smoke%20Detector%27,%20%27Building%20Dangerous%27)"


base311 = ("https://phl.carto.com/api/v2/sql?q=SELECT%20*%20FROM%20public_cases_fc%20WHERE%20")
where = paste("requested_datetime%20%3E=%20%27",Sys.Date()-730,
               "%27%20AND%20requested_datetime%20%3C%20%27", Sys.Date(),
               "%27%20AND%20service_name%20IN%20(%27Alley%20Light%20Outage%27,%20%27No%20Heat%20(Residential)%27%20,%20%27Fire%20Residential%20or%20Commercial%27%20,%20%27Infestation%20Residential%27%20,%20%20%27Smoke%20Detector%27,%20%27Building%20Dangerous%27)",sep="")

url311 <- paste(base311, where, sep="")

response311 <- httr::GET(url311)
tidy_res311 <- httr::content(response311, simplifyVector=TRUE)


if(length(tidy_res311$rows) != 0){
  request311 <- tidy_res311$rows %>%
    data.frame() %>%
    dplyr::select(service_request_id, status, service_name, service_code, requested_datetime, updated_datetime, address, lat, lon)
}else{
  request311 <- data.frame(Response=c("No relevant 311 request in the past year"))
}

request311.sf <- request311%>%
  drop_na(lat)%>%
  st_as_sf(coords = c("lon","lat"), crs = 4326)%>%
  st_transform(crs=3857)

#Nearest neighbor (NND) function
nn_function <- function(measureFrom,measureTo,k) {
  measureFrom_Matrix <- as.matrix(measureFrom)
  measureTo_Matrix <- as.matrix(measureTo)
  nn <-   
    get.knnx(measureTo, measureFrom, k)$nn.dist
    output <-
      as.data.frame(nn) %>%
      rownames_to_column(var = "thisPoint") %>%
      gather(points, point_distance, V1:ncol(.)) %>%
      arrange(as.numeric(thisPoint)) %>%
      group_by(thisPoint) %>%
      summarize(pointDistance = mean(point_distance)) %>%
      arrange(as.numeric(thisPoint)) %>% 
      dplyr::select(-thisPoint) %>%
      pull()
    
  return(output)  
}

light.sf<-request311.sf %>% 
  filter(service_name == 'Alley Light Outage')
heat.sf<-request311.sf %>% 
  filter(service_name == 'No Heat (Residential)') 
infestation.sf<- request311.sf %>%
  filter(service_name == 'Infestation Residential')
Detector.sf<- request311.sf %>%
  filter(service_name == 'Smoke Detector')
Dangerous.sf<- request311.sf %>%
  filter(service_name == 'Building Dangerous') 
  


DOR_Parcel.sf <- DOR_Parcel.sf %>%
  mutate(light.nn5 =  nn_function(st_coordinates(DOR_Parcel.sf),st_coordinates(light.sf), 5),
         heat.nn5 = nn_function(st_coordinates(DOR_Parcel.sf),st_coordinates(heat.sf), 5),
         infestation.nn5=nn_function(st_coordinates(DOR_Parcel.sf),st_coordinates(infestation.sf), 5),
         Detector.nn5 =nn_function(st_coordinates(DOR_Parcel.sf),st_coordinates(Detector.sf), 5),
         Dangerous.nn5 =nn_function(st_coordinates(DOR_Parcel.sf),st_coordinates(Dangerous.sf), 5),
         )
```


### 3.5 Census data

#### Load Census API Key
```{r load_api_key}
census_api_key("bdc91afe8f1e229bfd29314d696345b8365818b6", overwrite = TRUE)
```


#### Get Census Data
```{r tracts17}
dat<- 
  get_acs(geography = "tract",
          variables = c(pop="B01003_001",
                        whitepop="B01001A_001",
                        medinc="B06011_001",
                        blackpop = "B02001_003"),
          year=2019,
          state = "PA", 
          geometry = TRUE, 
          county="Philadelphia",
          output = "wide")

dat.sf <- dat %>%
  st_transform( crs = 3857)

census.sf <- 
  st_join(DOR_Parcel.sf,dat.sf,
          join=st_intersects,
          left = TRUE,
          largest = FALSE)

mapview(census.sf)
```

## 4 Nearby data

### 4.1 Request nearby properties data (including opa_account_num & Parcel_Id)
```{r Nearby opa&parcel_id, include=TRUE, echo=TRUE, message=FALSE}
ref_near_prop_url <- "SELECT * FROM opa_properties_public WHERE ST_DWithin(the_geom::geography,
ST_GeographyFromText('POINT(-75.1652 39.9526)'), 5)"

ref_near_prop_url <- "https://phl.carto.com/api/v2/sql?q=SELECT%20*%20FROM%20opa_properties_public%20WHERE%20ST_DWithin(the_geom::geography,%20ST_GeographyFromText(%27POINT(-75.1652%2039.9526)%27),%20100)"



base_near_prop <- "https://phl.carto.com/api/v2/sql?q=SELECT%20*%20FROM%20opa_properties_public%20WHERE%20ST_DWithin(the_geom::geography,%20ST_GeographyFromText(%27POINT("
LNG<- DOR_latlng$LNG[3]
LAT<- DOR_latlng$LAT[3]
end_near_prop <- ")%27),%2050)"
near_prop_url <- paste(base_near_prop, LNG,"%20" ,LAT, end_near_prop, sep="")

response_near_prop <- httr::GET(near_prop_url)
tidy_res_near_prop <- httr::content(response_near_prop, simplifyVector=TRUE)

near_prop <- tidy_res_near_prop$rows %>%
    data.frame() %>%
    dplyr::select(cartodb_id, parcel_number, registry_number, total_area, year_built, zoning, category_code, interior_condition) %>%
    mutate(category = case_when(category_code == 1 ~ "Residential",
                          category_code == 2 ~ "Hotels and Apartments",
                          category_code == 3 ~ "Store with Dwelling",
                          category_code == 4 ~ "Commercial",
                          category_code == 5 ~ "Industrial",
                          category_code == 6 ~ "Vacant Land"),
    interior = case_when(interior_condition == 0 ~ "Not Applicable",
                          interior_condition == 2 ~ "New/Rehabbed",
                          interior_condition == 3 ~ "Above Average",
                          interior_condition == 4 ~ "Average",
                          interior_condition == 5 ~ "Below Average",
                          interior_condition == 6 ~ "Vacant",
                          interior_condition == 7 ~ "Sealed/Structurally Compromised")) %>%
    rename(Parcel_Id = registry_number, opa_account_num = parcel_number) 
  
```

### 4.2 Request nearby parcels' geometry
```{r Nearby geometry, include=TRUE, echo=TRUE, message=FALSE}
for (i in 1:nrow(near_prop)) {
  if(near_prop$Parcel_Id[[i]] != "0LENGTH" && near_prop$Parcel_Id[[i]] != "NONE FOUND" &&near_prop$Parcel_Id[[i]] != "Parcel_Id IS NULL"){
      base <- "https://services.arcgis.com/fLeGjb7u4uXqeF9q/arcgis/rest/services/DOR_Parcel/FeatureServer/0/query?outFields=*&where=BASEREG%3D%27"
      BASEREG <- near_prop$Parcel_Id[[i]]
      end <- "%27&returnCentroid=true&f=pjson"
      dor_url <- paste(base, BASEREG, end, sep="")
      get_dor <- httr::GET(dor_url)
      dor_data <- httr::content(get_dor, simplifyVector=TRUE, as = "parsed", type = "application/json")
      near_prop[i,"Parcel_OBJECTID"] <- dor_data$features$attributes$OBJECTID
      near_prop[i,"PARCEL"] <- dor_data$features$attributes$PARCEL
      near_prop[i,"Shape__Area"] <- dor_data$features$attributes$Shape__Area
      near_prop[i,"Shape__Length"] <- dor_data$features$attributes$Shape__Length
      near_prop[i,"ADDR_SOURCE"] <- dor_data$features$attributes$ADDR_SOURCE
      near_prop[i,"x"] <- dor_data$features$centroid$x
      near_prop[i,"y"] <- dor_data$features$centroid$y
      #near_prop[i,"geometry"] <- dor_data$features$geometry
  }else{
      near_prop[i,"Parcel_OBJECTID"] <- "N/A"
      near_prop[i,"PARCEL"] <- "N/A"
      near_prop[i,"Shape__Area"] <- "N/A"
      near_prop[i,"Shape__Length"] <- "N/A"
      near_prop[i,"ADDR_SOURCE"] <- "N/A"
      near_prop[i,"x"] <- "N/A"
      near_prop[i,"y"] <- "N/A"
  }

}

near_prop.sf <- near_prop %>%
  drop_na(x)%>%
  st_as_sf(coords = c("x","y"), crs = 3857)

```

### 4.3 Nearby L&I violation

```{r L&I violation, include=TRUE, echo=TRUE, message=FALSE}
for (i in 1:nrow(near_prop.sf)) {
  base_url <- "https://phl.carto.com/api/v2/"
  endpoint <- "sql"
  query    <- c("?q=SELECT%20*%20FROM%20violations%20WHERE%20opa_account_num%20=%20")
  opa_num  <- paste0("%27",near_prop.sf$opa_account_num[[i]],"%27")
  url <- paste(base_url, endpoint, query, opa_num, sep="")
  response <- httr::GET(url)
  tidy_res <- httr::content(response, simplifyVector=TRUE)
  
  if (response$status_code != 400){
    if(length(tidy_res$rows$violationcode)==1){
    vio_code <-  tidy_res$rows$violationcode
    vio_title <- tidy_res$rows$violationcodetitle
  
    near_prop.sf$vio_code[[i]] <- vio_code
    near_prop.sf$vio_title[[i]] <- vio_title
    
    cat("Address",i,vio_code, vio_title, "\n")}
    else{
    near_prop.sf$vio_code[[i]] <- "NO CODE VIOLATION"
    near_prop.sf$vio_title[[i]] <- "NO CODE VIOLATION"
    cat("Address",i,"NO CODE VIOLATION\n")
    }
  }
  else{
    near_prop.sf$vio_code[[i]] <- "NO RESPONSE"
    near_prop.sf$vio_title[[i]] <- "NO RESPONSE"
    cat("Address",i,"NO RESPONSE\n")
  }
}
```

### 4.4 Nearby 311 request
```{r specific 311 request, include=TRUE, echo=TRUE, message=FALSE}
near_prop.sf <- near_prop.sf %>%
  mutate(light.nn5 =  nn_function(st_coordinates(near_prop.sf),st_coordinates(light.sf), 5),
         heat.nn5 = nn_function(st_coordinates(near_prop.sf),st_coordinates(heat.sf), 5),
         infestation.nn5=nn_function(st_coordinates(near_prop.sf),st_coordinates(infestation.sf), 5),
         Detector.nn5 =nn_function(st_coordinates(near_prop.sf),st_coordinates(Detector.sf), 5),
         Dangerous.nn5 =nn_function(st_coordinates(near_prop.sf),st_coordinates(Dangerous.sf), 5),
         )
```


### 4.5 Nearby census data
```{r Nearby census data, include=TRUE, echo=TRUE, message=FALSE}
near_prop.sf <- 
  st_join(near_prop.sf,dat.sf,
          join=st_intersects,
          left = TRUE,
          largest = FALSE)
```

